---
title: 'Data Analysis for: Priming Power with 3 DVs'
author: "Erika Salomon"
date: "January 12, 2016"
output: pdf_document
---

```{r, include=FALSE}
## Load Packages
library(scrapeR)
library(XML)
library(psych)
library(ggplot2)
library(GGally)
library(plyr)
library(rjson)
```

## Summary

This file reports data manipulations and analyses performed on an experiment studying the effects of power primes. Participants recruited via Mechanical Turk completed a power priming manipulation in which they were asked recall and visualize a time when they were high or low in power. After writing a description of their visualization, they completed a series of tasks to assess three dependent variables.

Participants'  **action/inaction intentions** were measured with 7 questions (e.g., "I am feeling energetic"; 1 = *not at all*, 7 = *extremely*; Albarracin, 2009, unpublished). 

Participants' **trust** was measured with two measures. Six items from the World Values Survey asked how much participants trust people belonging to different groups (e.g., "People you meet for the first time"; 1 = *trust completely*, 4 = *do not trust at all*). Participants also completed the trust game (Berg et al., 1995), which asked them how many raffle tickets they would entrust to another person who may or may not return them.

Two measures assessed participants' **perspective taking**. Participants answered seven items from the perspective-taking subscale of the Interpersonal Reactivity Index (Davis, 1980). These items were slightly modified to assess state rather than trait perspective taking motivation (e.g., "Before criticizing somebody, I'd try to imagine how I would feel if I were in their place"; 1 = *does not describe me at all*, 5 = *describes me very well*). Participants also completed 10 items from the Adult Reading the Mind in the Eyes Test-Revised (Baron-Cohen, Wheelwright, Hill, Raste & Plumb, 2001), which asked them to look at images of people's eyes and choose which of four emotions was expressed.

Participants completed measures for each of the dependent variables in a random order. For trust and perspective-taking, the two measures were also randomly ordered.

## Load the raw data

The data were collected using Qualtrics. The first row of the raw data file contains most of the variable names. The second row contains some further variable names and the question wording. Rows three and beyond contain participant data. So we will first load in the two sets of variable names and combine them. Then, we will load in the data.

```{r, cache=FALSE}
filename <- "PR__Priming_Power__3_DVs__Action_PerspectiveTaking_Distrust.csv"

name1 <- names(read.csv(filename, nrow = 1))
name2 <- names(read.csv(filename, nrow = 1, skip = 1))
names <- c(name2[1:22], name1[23:(length(name1))])
raw <- read.csv(filename, header = FALSE, skip = 2, col.names = names,
                stringsAsFactors = FALSE)
```

## Reduce data set to a final, usable set of cases

The final data set should only contain cases that represent:

+ US participants
+ Participants who passed the instructional manipulation check (IMC)

We will identify and count cases that do not meet these criteria. 

```{r}
# how many cases total?
totalCases <- nrow(raw)

# identify surveys from outside the US
nonUScount <- nrow(raw[with(raw, Country != "United States"),])

# identify participants who failed the IMC
failedIMCcount <- nrow(raw[with(raw, IMC != 10),])

# identify participants who meet both exclusion criteria
bothExcludeCount <- nrow(raw[with(raw, IMC != 10 & Country != "United States"),])

# create final data set and count cases
final <- raw[with(raw, IMC == 10 & Country == "United States"),]
finalCases <- nrow(final)
```

From `r totalCases` raw cases, `r nonUScount` had non-US IP addresses and `r failedIMCcount` failed the IMC (`r bothExcludeCount` failed the IMC *and* had a non-US IP), leaving `r finalCases` for analysis.

## Create labels for categorical/ordinal demographic variables

The categorical and ordinal demographic variables (e.g., race, gender) are coded with numbers. We want to add category labels to these. 

```{r}
# ethnicity/race
final$ethnic2 <- factor(final$ethnic2,
                        levels = c(1,2,3,4,5,6),
                        labels = c("African American",
                                   "Asian American",
                                   "European American",
                                   "Latino/Latina",
                                   "Native American",
                                   "Other"))

# gender
final$gender <- factor(final$gender,
                       levels = c(1,2),
                       labels = c("Male", "Female"))

# income 
final$yyour <- factor(final$yyour,
                      levels = c(1,2,3,4,5,6,7,8),
                      labels = c("Under $15,000",
                                 "$15,001 - $25,000",
                                 "$25,000 - $35,000",
                                 "$35,001 - $50,000",
                                 "$50,001 - $75,000",
                                 "$75,001 - $100,000",
                                 "$100,001 - $150,000",
                                 "Over $150,000"))

# education 
final$youred <- factor(final$youred,
                       levels = c(1,2,3,4,5),
                       labels = c("Less than high school",
                                  "High school graduation or equivalent",
                                  "Some college", "College graduation",
                                  "Professional / Post-graduate degree"))
```

## Create condition variables for analyses

Conidition was recorded with a dummy code (0 = Low Power; 1 = High Power). We will convert these to effect codes with a 1 unit difference, weighted by the number of participants randomly assigned to the other condition. Weighted effect codes will produce regression coefficients that represent the mean difference between the groups. 

```{r}
# how many participants were assigned to each condition?
lowN <- nrow(final[final$Cond == 0,])
highN <- nrow(final[final$Cond == 1,])

# create weights
final$effect[final$Cond == 0] <- -highN / finalCases
final$effect[final$Cond == 1] <- lowN / finalCases
```

There were `r lowN` and `r highN` participants in the Low and High Power conditions, respectively.

We will also turn the original dummy code into a factor, which will be useful for plotting and tables.

```{r}
final$Cond <- factor(final$Cond,
                     levels = c(0,1),
                     labels = c("Low Power", "High Power"))
```


## Exploratory data analysis

Before we run our preferred model, we should use exploratory data analysis to examine the distributions of variables and investigate the potential influence of order effects on the dependent variables.

### Descriptive summaries of demographic variables

First, we will look at the demographic composition of our sample.
 
```{r}
# Gender
table(final[,"gender"])

# Age
describe(final[,"age"])

# Race/ethnicity
prop.table(table(final$ethnic2))

# Income
prop.table(table(final[,"yyour"]))

# Education
prop.table(table(final$youred))
```

Overall, the sample looks similar to other Mechanical Turk samples.
